---
title: "PE Study"
author: "Carlos Brown"
date: "9/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(exact2x2)
library(tidyr)
library(dlookr)
library(exact2x2)
library(ISLR)
library(VGAM)
library(ggplot2)
library(samplesizeCMH)
library(DescTools)

# file_saddle <- file.path("saddle.csv")
# file_nosaddle <- file.path("nosaddle.csv")
file_all <- file.path('all_types.csv')

# df_saddle <- read.csv(file_saddle, header = TRUE)
# df_nosaddle <- read.csv(file_nosaddle, header = TRUE)
df_model <- read.csv(file_all, header = TRUE)
```

Within this study, the term adverse event rate and adverse event count are synonymous.

## Initial Processing

Change types and rename
```{r change type to factor and rename}
df_model$type <- as.factor(df_model$type)
df_model <- df_model %>% rename(Severity = PE.Severity.1.low.risk..2.intermediate..3.massive..4.indeterminate)
df_model$Gender <- as.factor(df_model$Gender)
df_model$Race <-as.factor(df_model$Race)
df_model$Severity <- as.factor(df_model$Severity)
```

Filter for low risk patients by sPESI
```{r Initial filter for low risk patients based on sPESI }
df_all <- df_model %>% filter(sPESI == 0)
```

Bin by age
```{r bin by age}
breaks = 5
df_all$age_bin <- cut(df_all$Age, breaks=breaks)
```

## What constitutes an Adverse Event?
1. Death, either in hospital or within 30 days of release
2. Arrhythmia that required treatment
3. Inotropes
4. Vasopressors
5. Use of Non-Rebreather mask
6. ICU Stay
7. Shock
8. ACLS
9. Intubation
10. tPA
11. Transfusion
12. EKOS

## Visualize Distributions
```{r ggplot barchart, echo=FALSE}
df_plot <- df_all %>% mutate(adverseEvent = as.factor(adverseEvent))
g_plot <- ggplot(data=df_plot)
g_plot + geom_bar(mapping = aes(x=type, fill=adverseEvent)) + ggtitle('Adverse Event Count by PE Type')
```

```{r adverse event count by PE severity, echo=FALSE}
g_plot <- ggplot(data=df_plot)
g_plot + geom_bar(mapping = aes(x=Severity, fill=adverseEvent)) + ggtitle('Adverse Event Count by PE Severity')
```

```{r Adverse Events by Age Range, echo=FALSE}
g_plot <- ggplot(data=df_plot)
g_plot + geom_bar(mapping = aes(x=age_bin, fill=adverseEvent)) + ggtitle('Adverse Event Count vs Age Range')
```

```{r Severity by PE Type, echo=FALSE}
g_plot <- ggplot(data=df_plot)
g_plot + geom_bar(mapping = aes(x=type, fill=Severity), position = "fill") + ggtitle('Severity by PE Type')
```

Severity correlates highly with pe type, therefore is a potential confounder

## Check for confounding
```{r test whether confounding is present, crude model}
m_crude <- glm(formula = 'adverseEvent ~ type', data = df_all, family = binomial)
summary(m_crude)
```

```{r test whether confounding is present, adjusted}
m_adj <- glm(formula = 'adverseEvent ~ type + Age + Severity + Race + Gender', data = df_all, family = binomial)
summary(m_adj)
```

Calculate percent change in typesaddle coefficient between models. If absolute value is greater than 10%, confounding needs to be adjusted for.
```{r percent change}
beta_crude <- coefficients(m_crude)[['typesaddle']]
beta_adj <- coefficients(m_adj)[['typesaddle']]
(beta_adj - beta_crude) / beta_crude
```
Confounding is present.

Compare to VGAM package outputs for sanity check
```{r compare to VGAM for sanity check}
m_vgam <- vglm(formula = 'adverseEvent ~ type + Age + Severity + Race + Gender', binomialff(link = "logitlink"), data = df_all)
summary(m_vgam)
```

The predictor "type" is confounded by adding in further predictors such as Age, Severity, and Race. Severity alone is the most powerful confounder and should be controlled for. Age itself is not a significant confounder. We will use the Cochran-Mantel-Haenszel to account for the observed confounding effect of Severity, of which only Severities of 1 and 2 have enough sample points to be counted.

## 2x2 Tables

2x2 Table for entire cohort
```{r 2x2 in just 2 dimensions}
# table_group <- table(df_all$adverseEvent, df_all$type, df_all$age_bin)
test_mat <- as.matrix(table(df_all$type, df_all$adverseEvent))
test_mat
```


2x2 Table for cohort, by Severity
```{r 2x2 in 3 dimensions}
table(df_all$type, df_all$adverseEvent, df_all$Severity)
```


Hypothesis test is set based on what constitutes surprise for the doctor.

H0: There is no difference in the adverse event odds ratio of saddle vs non-saddle in age and severity controlled groups

H1: There is a difference in the adverse event odds ratio of saddle vs non-saddle in age and severity controlled groups

Odds ratio in this case is defined as...

(saddle adverse event count / saddle no adverse event count) / (non-saddle adverse event count / non-saddle no adverse event count)

Test 1: Fisher's exact test for entire cohort
```{r Run Fishers exact test for entire set}
m_fisher <- exact2x2(test_mat, conf.level = 0.99)
m_fisher
```

There is evidence to reject the null hypothesis, however we have to control for confounding factor of PE Severity. Use the Cochran-Mantel-Haenszel test.

```{r Cochran-Mantel-Haenszel to control for Severity}
m_cmh <- mantelhaen.test(table(df_all$type, df_all$adverseEvent, df_all$Severity), exact = TRUE)
m_cmh
```

We notice that after controlling for PE Severity in low risk patients, there is still evidence to suggest a statistically significant and practically significant difference in the adverse event counts between saddle and non-saddle pulmonary embollism.

Control for multiple comparisons using Benjamini & Yekutieli correction
```{r control False Discovery Rate using BH or BY}
pvalues <- c(m_fisher$p.value, m_cmh$p.value)
p.adjust(pvalues, method = "BY")
```

Given the multiple comparisons correction, and our alpha value of 1%, both tests have given p-values where we can reject the null hypothesis in favor of the alternate hypothesis. The common odds ratio confidence interval for adverse events in saddle over non-saddle is...

`r m_cmh$conf.int`

Patients with Saddle PE are `r round(m_cmh$estimate, digits = 2)` times as likely to experience an adverse event as those with Non-Saddle PE


## How predictive are PESI and sPESI?

```{r predictive model sPESI}
m_pred_spesi <- glm(formula = 'adverseEvent ~ sPESI', data = df_model, family = binomial)
PseudoR2(m_pred_spesi, which = 'McKelveyZavoina')
summary(m_pred_spesi)
```

```{r predictive model PESI}
m_pred_pesi <- glm(formula = 'adverseEvent ~ PESI.Class', data = df_model, family = binomial)
PseudoR2(m_pred_pesi, which = 'McKelveyZavoina')
summary(m_pred_pesi)
```

Both sPESI and PESI Class are significant predictors of adverse event rates and have similar predictive power.

```{python logit, eval=FALSE, include = FALSE}
import statsmodels.discrete.discrete_model as sm

feat_cols = ['Age','Hight-cm','Weight-kg','HR','SBP','Temp (F)','PA/Aorta Ratio','RV/LV Ratio', 'PESI.Class' , 'sPESI', 'Race', 'adverseEvent']

features = df_saddle[feat_cols].dropna()
target = features['adverseEvent']

features.drop(columns='adverseEvent', inplace=True)
logit = sm.Logit(target.values, features.astype(float))
res = logit.fit()

print(res.summary())
```

```{r control for age, eval=FALSE, include=FALSE}
age_vector <- levels(df_all$age_bin)

evaluate_fisher <- function(bin, severity){
  df_bin <- df_all %>% filter(age_bin == get("bin"))
  severities <- unique(df_bin$Severity)
  for (sev in severities) {
    model_obj <- exact2x2(as.matrix(table(df_bin$adverseEvent, df_bin$type)), conf.level = 0.99)
  }
  
  return(model_obj)
}
age_controlled_results <- lapply(age_vector, evaluate_fisher)

```
